---
title: "Moving averages in R"
---


```{r}
# setup
install.packages("slider")
library(fpp3)
```


## Defining Moving average

A moving average of order $m$, with $m = 2k + 1$, is defined as: 
$$\hat T_t = \frac{1}{m} \sum_{j = -k}^{k} y_{t + j}$$

Where:

- $\hat T_t$ is the estimated moving average at time $t$.
- $m$ is the order of the moving average. 
- $y_{t + j}$ is the value of the time series at time $t + j$.
- $k$ is the number of observations to average on either side of time $t$.


```{r}
# from fpp3 text
global_economy |>
  filter(Country == "Australia") |>
  autoplot(Exports) +
  labs(y = "% of GDP", title = "Total Australian exports")
```





### Add moving average to plot

```{r}
aus_exports <- global_economy |>
  filter(Country == "Australia") |>
  mutate(
    `5-MA` = slider::slide_dbl(Exports, mean,
                .before = 2, .after = 2, .complete = TRUE)
  )
```

```{r}
# examine the data frame
aus_exports
```

```{r}
aus_exports |>
  autoplot(Exports) +
  geom_line(aes(y = `5-MA`), colour = "#D55E00") +
  labs(y = "% of GDP",
       title = "Total Australian exports")
```

## Question

Write, in terms of the original observations, the order 5 moving average at time point 3.






## Moving average of moving averages

- **Why?** Might want to work with *even* order moving averages, but need to make them symmetric    
- eg., Estimating the trend-cycle from seasonal data  


```{r}
beer <- aus_production |>
  filter(year(Quarter) >= 1992) |>
  select(Quarter, Beer)
beer_ma <- beer |>
  mutate(
    `4-MA` = slider::slide_dbl(Beer, mean,
                .before = 1, .after = 2, .complete = TRUE),
    `2x4-MA` = slider::slide_dbl(`4-MA`, mean,
                .before = 1, .after = 0, .complete = TRUE)
  )
```


```{r}
# look at the resulting data frame 
beer_ma
```

## Questions


Question: 

Write, in terms of the original observations, the order 5 moving average at time point 3.



Question: 

Write, in terms of the original observations, the order 2 moving average of the order 4 moving average at time point 3.





## Estimating the trend-cycle from seasonal data


```{r}
us_retail_employment <- us_employment |>
  filter(year(Month) >= 1990, Title == "Retail Trade") |>
  select(-Series_ID)
us_retail_employment_ma <- us_retail_employment |>
  mutate(
    `12-MA` = slider::slide_dbl(Employed, mean,
                .before = 5, .after = 6, .complete = TRUE),
    `2x12-MA` = slider::slide_dbl(`12-MA`, mean,
                .before = 1, .after = 0, .complete = TRUE)
  )
```

```{r}
# look at the us_retail_employment_ma
us_retail_employment_ma
```


```{r}
# plot without moving averages
us_retail_employment_ma |>
  autoplot(Employed, colour = "gray") +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
```



```{r}
# plot with moving averages
us_retail_employment_ma |>
  autoplot(Employed, colour = "gray") +
  geom_line(aes(y = `2x12-MA`), colour = "#D55E00") +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
```


